---
title: "Homework 2 - SDS"
author: "Andrea Conti - Davide Mascolo - Mario Napoli"
date: "27 dicembre 2021"
output: html_document
---

### 1. Introductory
Let's define the remaining $\nu$ values:
$\nu(AB) = 2, \quad \nu(AC) = 0, \quad \nu(BC) = 2$.
Now we can build the table as follows:

Permutations | A | B | C
:----: | :----: | :----: | :----:
ABC  | 0 | 2 | 2
ACB  | 0 | 4 | 0
BAC  | 2 | 0 | 2
BCA  | 2 | 0 | 2
CAB  | 0 | 4 | 0
CBA  | 2 | 2 | 0

From this table we can compute Shapley values:
$\psi(A) = \frac{6}{6} = 1, \quad \psi(B) = \frac{12}{6} = 2, \quad \psi(C) = \frac{6}{6} = 1$.
We can normalize these values in a way that they sum up to 1, in other words we can divide them by 4 but the interpretation would remain the same: A save 1 hour of work (or 25% of the total savings if we normalize), B 2 hours (or 50% of the total) and C another hour (the remaining 25%).

### 2. Probabilistic
```{r, results=F, message=F}
## Utils
library(tseries)
library(igraph)
library(knitr)
```

```{r}
## Load CharFun
load('char_fun.RData')
```

```{r}
options(scipen = 999)
```

We separate coalitions in which it is present *j* from those in which *j* is not present.
```{r}
## Split lists
split_list <- function(j){
  # Input: a specific player
  # Output: a list with all the coalitions in which j is not present
  no_j <- vector(mode = "list", length = 2048-1)
  k <- 1
  for (i in 1:12){  
      for (c in 1:length(char_fun[[i]])){
        name <- names(char_fun[[i]][c])
        value <- char_fun[[i]][[c]]
        if (!(j %in% as.integer(strsplit(name, '-')[[1]]))){
          no_j[[k]] <- c(name, value)
          k = k + 1
        }
      }
  }
  return(no_j)
}
```

Compute the differences between the values of the coalitions without the player *j* and the coalitions with the player *j*.
```{r}
compute_diff <- function(j){
  # Input: a specific player
  # Output: list in which are stored the difference: C U {j} - C
  no_j <- split_list(j)
  diff_list <-c()
  for (i in 1:length((no_j))){
    coal <- no_j[[i]][1]
    val_coal <- as.integer(no_j[[i]][2])
    # Just elaboration to obtain the number (players) in coalition C U {j}
    coal2 <- paste(sort(c(as.numeric(unlist(strsplit(coal, split = '-'))), j)), collapse = '-')
    tmp <- strsplit(coal2, split = '-')[[1]]
    val_coal2 <- char_fun[[length(tmp)]][coal2][[1]]
    diff_list <- c(diff_list, val_coal2 - val_coal)
  }
  return(diff_list)
}
```

```{r}
## Simulation
simul <- function(j){
  # Input: a specific player
  # Output: our estimated Shapley values plus more things useful for later steps:
  #         - max and min of the difference list
  #         - M, our computed simulation size
  alpha <- 0.05
  epsilon <- 0.1
  
  diff_list <- compute_diff(j)
  
  M <- ceiling(((max(diff_list) - min(diff_list))^2*log(2/alpha))/(2*(epsilon^2)))
  tot <- 0
  
  for (i in 1:M){
    n <-sample(diff_list, 1, replace = TRUE)
    tot = tot + n
  }
  
  return(c(tot / M, max(diff_list), min(diff_list), M))
}
```

We run the simulation for the three selected players and compute the respective confidence intervals.
```{r}
## Set players
players <- c(2, 3, 5)

df <- data.frame(matrix(0, nrow = 3, ncol = 3))
colnames(df) <- c('LowerBound', 'Value', 'UpperBound')
rownames(df) <- c(players)
i <- 1
for (p in players){
  shap_ret = simul(p)
  shapley_val = shap_ret[1]
  mx <- shap_ret[2]
  mn <- shap_ret[3]
  M <- shap_ret[4]
  eps_alpha <- sqrt(((mx - mn)^2 * log(2/0.05)) / (2*M))
  df$LowerBound[i] <- shapley_val - eps_alpha
  df$Value[i] <- shapley_val
  df$UpperBound[i] <- shapley_val + eps_alpha
  i = i + 1
}
kable(df)
```

We set $\alpha = 0.05$ and $\epsilon = 0.1$.
Then, starting from the following formula $\epsilon_\alpha = \sqrt{\frac{(b-a)^2}{2M}\log{(\frac{2}{\alpha})}}$ we reversed it in order to compute $M$, that is the size of the simulation that allows us to reach a $1-\alpha$ confidence interval. 
After that, we computed the confidence intervals of the Shapley values for each player.
Regardless of the value of $M$, we can note that $\epsilon_\alpha$ is almost the same for the three players that we have chosen. This implies that the confidence intervals for the three different players about the same width due to our "$\alpha$-level".

### 3. Statistical
##### List of S&P 500

* AMZN - Customer Discretionary
* MSFT -  Informational Technlogy
* BLK  - Financials
* CAH  - Healt-Care
* VZ   - Communication Services
* CVX  - Energy
* ECL  - Materials
* EIX  - Utilitis
* DAL  - Industrials
* SYY  - Customer Staples 

```{r}
## Load Data Function
load_data <- function(code){
  # Input: a specific code of stock (e.g. Amazon -> AMZN)
  # Output: the historical series of that title
  
  data <- get.hist.quote(instrument = code,
                         start = '2020-01-01',
                         end = '2021-12-22',
                         quote = c('Open', 'Close'),
                         provider = 'yahoo',
                         drop = TRUE)
  return(data)
}
```

Load data
```{r, results=F, message=F}
options("getSymbols.warning4.0"=FALSE)
## AMZN
amz_data <- load_data('AMZN')
## MSFT
msft_data <- load_data('MSFT')
## BR
blk_data  <- load_data('BR')
## CAH
cah_data  <- load_data('CAH')
## VZ
vz_data   <- load_data('VZ')
## CVX
cvx_data  <- load_data('CVX')
## ECL
ecl_data  <- load_data('ECL')
## EIX
eix_data  <- load_data('EIX')
## DAL
dal_data  <- load_data('DAL')
## SYY
syy_data  <- load_data('SYY')
```

Build portfolio
```{r}
port <- data.frame(matrix(0, nrow(amz_data-1), 10))
names(port) <- c('AMZN', 'MSFT', 'BLK', 'CAH', 'VZ',
                 'CVX', 'ECL', 'EIX', 'DAL', 'SYY')

## Populate Port
for (i in (2:nrow(amz_data))){
  port$AMZN[i-1]<-log(amz_data$Close[[i]]/
                        amz_data$Close[[i-1]])
  port$MSFT[i-1]<-log(msft_data$Close[[i]]/
                      msft_data$Close[[i-1]])
  port$BLK[i-1]<-log(blk_data$Close[[i]]/
                       blk_data$Close[[i-1]])
  port$CAH[i-1]<-log(cah_data$Close[[i]]/
                       cah_data$Close[[i-1]])
  port$VZ[i-1]<-log(vz_data$Close[[i]]/
                       vz_data$Close[[i-1]])
  port$CVX[i-1]<-log(cvx_data$Close[[i]]/
                        cvx_data$Close[[i-1]])
  port$ECL[i-1]<-log(ecl_data$Close[[i]]/
                        ecl_data$Close[[i-1]])
  port$EIX[i-1]<-log(eix_data$Close[[i]]/
                        eix_data$Close[[i-1]])
  port$DAL[i-1]<-log(dal_data$Close[[i]]/
                        dal_data$Close[[i-1]])
  port$SYY[i-1]<-log(syy_data$Close[[i]]/
                        syy_data$Close[[i-1]])
}

## View
kable(head(port))
```

```{r}
## Estimate Correlation Matrix
kable(cor(port), allign = 'cccccccccc')

## List of Companies
companies <- c(names(port))

## Init Graph (Undirected)
net <- make_empty_graph(10, F)  

## Vertex names
V(net)$name <- companies

## Colors
colors <- c('gray', 'green', 'blue', 'red')
```

We place an edge only if 0 doesn't belong in the confidence interval and we choose the color of this edge based on the correlation value.
```{r}
for (i in 1:(length(companies)-1)){
  for (j in (i+1):length(companies)){
    cor_value <- cor.test(port[,i],
                          port[,j])
    conf_int <- cor_value$conf.int
    est_cor  <- cor_value$estimate
    if (!((conf_int[1]<= 0) & (conf_int[2]>=0))){
      if (est_cor <= 0.25)
        col <- colors[1]
      else if (est_cor <= 0.50)
        col <- colors[2]
      else if (est_cor <= 0.75)
        col <- colors[3]
      else
        col <- colors[4]
      net <- add_edges(net, c(companies[i],
                              companies[j]),
                       color = col)
    }
  }
}
```

```{r, fig.height = 10, fig.width = 10,fig.align = 'center'}
## Plot Graph
plot(net)
legend('bottomleft',
       legend = c('[0:0.25]', '(0.25:0.50]',
                  '(0.50:0.75]', '(0.75:1]'),
       title = 'Correlation Bounds',
       lty = 1,
       col = colors)
```
We choose to pick one company from each *GICS Sector* in order to diversificate our portfolio, this is also an attempt to reduce correlations (e.g. we imagine that Facebook and Google are more co-related than Google and an energy company) between stocks.
We also choose a *95% Normal* confidence interval setting $\alpha=0.05$. With this setting no correlation contains 0 within the range and no correlation exceeds the value of 0.75 and this, in our graph, results in a complete graph, despite this we noticed that more than a few stocks are weakly correlated with others and this is the reason why we choose to place edges of different colors.

```{r}
## Estimate Shapley Values
sv <- rep(0, ncol(port))
sv_hat <- function(X, omega = 0.7){
  for (i in 1:ncol(port)){
    sv[i] <- mean(X[,i]) + sum(cov(X)[,i]) * omega
  }
  return(sv)
}

## Check
shapley <- sv_hat(port)
```

```{r, warning=F, comment=NA}
## Bootstrap
B <- 1000
sv_boot <- rep(NA, B)

## Cycle
for (i in 1:B){
  idx       <- sample(1:nrow(port), replace = T)
  port_boot <- port[idx, ]       ## Bootrap sample
  sv_boot[i]<- sv_hat(port_boot) ## Bootstrap Shapley Value
}

## Check
c(head(sv_boot / B, 10))
```

```{r}
## Bootstrap Standard Error
se_boot <- sqrt(var(sv_boot))

## Confidence Level
alpha   <- 0.05

## Bootstrap / Normal CI / 95%
z <- qnorm(1 - alpha / 2)

## Check
df <- data.frame(shapley - z * se_boot,
                 shapley,
                 shapley + z * se_boot)
## Rename columns
colnames(df) <- c('LowerBound', 'Value', 'UpperBound')
rownames(df) <- companies

## View
kable(df, allign = 'ccc')
```

Setting $\omega=0.7$, $\alpha=0.05$ and based on a *Normal Confidence Interval*, we estimate the *Shapley Values* for each stock with the respective CI, based on Bootstrap sampling. We have chosen this value of $\omega$ because we think that *expected return* of a stock is more important than the *volatility* that refers to a degree of fluctuation of the asset returns because we assume that when someone invests his money he is interesting in the expected revenue despite the possible fluctuations that the *title* may have.


We discussed about this homework with Cruoglio Antonella and Iovino Giuliana.